---
title: "20241024_Writing_Functions"
output: github_document
date: "2024-10-24"
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(rvest)
```

rvest to do some data importing 

Functions to use when using the same basic code more than two times (iteration)
Functions like mapping in maths (have an input then have a described operation and have an output)

Scoping: if in function there is some variable that you haven't defined then the function will scope your entire R environment to see whether you had previously defined this variable (but generally this is something that you do not want)
want to make sure that everything that you need for the function to work is defined inside the function (since work environemnt can change)
Function should be self contained and is everything that this function needs defined in the function 

Writing a function that is clear is really important 

## Writing my first function 

Standardising is frequent in statistic (subtract mean divide standard deviation)

As an example, here is a z-score computation 

```{r}
x_vec = rnorm(n = 25, mean = 10, sd = 3.5)

(x_vec - mean(x_vec)) / sd(x_vec)
```

Next write a function that calculates the z-score 

```{r}
z_scores = function(x) {
  
  z = (x - mean(x)) / sd(x)
  
  return(z)
}

z_scores(x = x_vec)
```

parenthesizes () is where define argument, {} to define the operation
  Can see () defining what my inout is and {} what my output should be 

Inside the function not re-using the x-vector 
Inside function i have the argument x and then outside the function i have defined that the x argument is the x-vector
  Can change the argument into something else and the function will operate the same 
  
Can create a new x vector 

Kinds of things that want to double check 
Does this always work

```{r, error = TRUE}
z_scores(x = 3)
```
Cannot work since there is no sd if there is only one value 

```{r, error = TRUE}
z_scores("my name is jeff")
```
Cannot work since trying to take the mean of words not numbers 


So these are checks and we can put this into the function 
Conditional execution checks
```{r}
z_scores = function(x) {
  
  if(!is.numeric(x)) {
    stop("x needs to be numeric")
  }
  
  if(length(x) < 5) {
    stop("you need at least 5 numbers to compute the z score")
  }
  
  z = (x - mean(x)) / sd(x)
  
  return(z)
}

z_scores(x = x_vec)
```

```{r, error = TRUE}
z_scores(x = 3)
```

```{r, error = TRUE}
z_scores("my name is jeff")
```

Now these can work

Helpful that conditional execution exists 
If the input is a character vector i want this to happen but if it is numeric i want this to happen (compute a plot or a table based on data input)

This is relatively simple (one input one output)

## A new function! Multiple outputs

```{r}
mean_and_sd = function(x) {
  
  mean_x = mean(x)
  sd_x = sd(x)
  
  out_df = 
    tibble(
      mean = mean_x, 
      sd = sd_x
    )
  
  return(out_df)
}

mean_and_sd(x_vec)
```

In this function () this is my input and my {} defines what i want to do and here i want to do two things, compute the mean and sd and i want the function to give me both results 

Rather than have it spit it out, format an output dataframe 


## Check stuff using a simulation 

Checking that the sample mean and sd are close to true population mean and sd 

Create a simulation df with one column
```{r}
sim_df = 
  tibble(
    x = rnorm(30, 10, 5)
  )

sim_df %>% 
  summarise(
    mean = mean(x),
    sd = sd(x)
  )
```

Could run this over and over to see different samples of 30 variables of the normal dis to see whether close to true mean 10 and true sd 5

```{r}
sim_mean_sd = function(samp_size, true_mean, true_sd) {
  
  sim_df = 
  tibble(
    x = rnorm(samp_size, true_mean, true_sd)
  )

out_df =
  sim_df %>% 
  summarise(
    mean = mean(x),
    sd = sd(x)
  )

return(out_df)

}

sim_mean_sd(samp_size = 3000, true_mean = 4, true_sd = 12)
```

Can play around here to see whether Central Limit Theorem works
Increasing sample size and smaller sd 

See that i defined my argument (samp_size = 3000, true_mean = 4, true_sd = 12), but we can take advantage that it will read values by order so could just write it as (30, 4, 12) at the end and it would still run the same 

```{r}
sim_mean_sd = function(samp_size, true_mean, true_sd) {
  
  sim_df = 
  tibble(
    x = rnorm(samp_size, true_mean, true_sd)
  )

out_df =
  sim_df %>% 
  summarise(
    mean = mean(x),
    sd = sd(x)
  )

return(out_df)

}

sim_mean_sd(3000, 4, 12)
```


If we want to set the true mean and true sd constant and just have sample size varied then can define in the start 

```{r}
sim_mean_sd = function(samp_size, true_mean = 5, true_sd = 6) {
  
  sim_df = 
  tibble(
    x = rnorm(samp_size, true_mean, true_sd)
  )

out_df =
  sim_df %>% 
  summarise(
    mean = mean(x),
    sd = sd(x)
  )

return(out_df)

}

sim_mean_sd(samp_size = 3000, true_mean = 4)
```

Just because there is a default value doesn't mean that you cannot overwrite it 

See above that we set default for true mean and true sd but then we overwrote the true mean and kept the default for true sd 

## LOTR dataframe

```{r}
fellowship_df = 
  read_excel("data/LotR_Words.xlsx", range = "B3:D6") %>% 
  mutate(movie = "fellowship")

two_towers = readxl::read_excel("data/LotR_Words.xlsx", range = "F3:H6") %>% 
  mutate(movie = "two_towers")

return_king = readxl::read_excel("data/LotR_Words.xlsx", range = "J3:L6") %>%
  mutate(movie = "return_king")
```

This is repetitive to import the LOTR data can be easy to make mistakes doign iteratively so create a function that does it 

Think about the steps that i took for each one: path, range, mutate movie name

```{r}
LOTR_import = function(cell_range, movie_title) {
  
  movie_df = 
    read_excel("data/LotR_Words.xlsx", range = cell_range) %>% 
    mutate(movie = movie_title) %>% 
    janitor::clean_names() %>% 
    pivot_longer(
      female:male,
      names_to = "sex",
      values_to = "words"
    ) %>% 
    select(movie, everything())
  
  return(movie_df)
}

lotr_df = 
  bind_rows(LOTR_import("B3:D6", "fellowship"),
            LOTR_import("F3:H6", "two_towers"),
            LOTR_import("J3:L6", "return_king")
  )
```

Functions has to have 3 things: inputs, operation, output 

Think about what i needed to change (cells names and movie title)
With the supplied cell range and movie title

Can put into the function operations a code chunk that you know already exists then substitute the specific variables with the general variable names To make them arguments rather than specific inputs 

Now can iteratively change all, updated with pivoting to pivot all and not have to write it out each time 


## NSDUH 

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

data_marj = 
  nsduh_html %>% 
  html_table() %>% 
  nth(1) %>% 
  slice(-1) %>% 
  mutate(drug = "marj")

data_coca = 
  nsduh_html %>% 
  html_table() %>% 
  nth(4) %>% 
  slice(-1) %>% 
  mutate(drug = "cocaine")

data_hero = 
  nsduh_html %>% 
  html_table() %>% 
  nth(5) %>% 
  slice(-1) %>% 
  mutate(drug = "heroin")

```

Keep track of what i am changing everytime (so that i can find the backbone of my function)
```{r}
NSDUH_import = function(table_num, drug_name) {
  
  drug_df = 
    nsduh_html %>% 
    html_table() %>% 
    nth(table_num) %>% 
    slice(-1) %>% 
    mutate(drug = drug_name)
  
  return(drug_df)
}
  
NSDUH_import(1, "marj")
NSDUH_import(4, "cocaine")
NSDUH_import(5, "heroin")

```

REALISE THAT FOR THIS THE FUNCTION IS SCOPING for nsduh_html since it is not defined in the input argument (table_num, drug_name)
it has scoped my R environment what nsduh_html is 

Can change this to define better the html 

```{r}
NSDUH_import = function(html, table_num, drug_name) {
  
  drug_df = 
    html %>% 
    html_table() %>% 
    nth(table_num) %>% 
    slice(-1) %>% 
    mutate(drug = drug_name)
  
  return(drug_df)
}
  
NSDUH_import(html = nsduh_html, 1, "marj")
NSDUH_import(html = nsduh_html, 4, "cocaine")
NSDUH_import(html = nsduh_html, 5, "heroin")

```

A stylistic choice is to keep a source folder and an R-script with the function there. 
This means that do to have to have the function there (do not need to necessarily see the function in the R Markdown file just need to see that it works, also useful if have multiple R Markdown files that want to use the same function then can just continue to source from R script rather than repeated function)

Writing a function to get each table
```{r}
source("functions_source/nsduh_table_format.R")
nsduh_df = bind_rows(
  NSDUH_import(html = nsduh_html, 1, "marj"),
  NSDUH_import(html = nsduh_html, 4, "cocaine"),
  NSDUH_import(html = nsduh_html, 5, "heroin")
)
```

