---
title: "20241029_Iteration_and_Listcols"
output: github_document
date: "2024-10-29"
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
```

We write a function because we would do the same basic code over and over again. Process doing over and over is iterations. Take advantage of the formal structure 

Loop functions (avoid keeping track input / output list), easier to read 

Map rather than a for loop (not looking inside a for loop to look for index, instead here is the function that i have and here is input / output)

Process:
1. start with single working example 
2. change into a function 
3. re-write into a map statement 
(do not immediately start try writing a mapping statement)

Listcols --> 
lists are just collections of things 
not a dataframe (completely arbitary into a list)
Dataframes are a fancy / specific list 
Lists can also contain lists (there are lots of stuff to keep track of, if can keep organised into datafram it would be really helpful, also sub-elements (there are 1000 weather station, where each weather station has its own collection of analyses), works well for this kind of iterative analysis) 

# Here's some lists

Can do something inside the list (define a vector)
Different sorts of objects that can all exist in a single list
```{r}
l = list(
  vec_numeric = 1:4, #1 - 4
  unif_sample = runif(100), #100 numbers 
  mat = matrix(1:8, nrow = 2, ncol = 4, byrow = TRUE),
  summary = summary(rnorm(1000))
)

l$mat
```

notation l$mat can pull out from list (not the most ideal with $ but can) or can l[["mat"]] with two square brackets

```{r}
l[["mat"]]

l[["mat"]][1, 3] #in mat list find either 

l[[1]] #pull the first object from list
```

```{r}
list_norm = 
  list(
    a = rnorm(20, 0, 5),
    b = rnorm(20, 4, 5),
    c = rnorm(20, 0, 10),
    d = rnorm(20, 4, 10)
  )

list_norm[["b"]] #can pull specific one
#list of 4 samples from 4 differnet normal dis
```

Now have thees 4 samples and want to compute the mean and sd of each of the norm dis (write a function for mean and standard dev)

```{r}
mean_and_sd = function(x) {
  
  mean_x = mean(x)
  sd_x = sd(x)
  
  out_df = 
    tibble(
      mean = mean_x, 
      sd = sd_x
    )
  
  return(out_df)
}

```

We can use this function to make mean and sd of all samples

```{r}
mean_and_sd(list_norm[["a"]])
mean_and_sd(list_norm[["b"]])
mean_and_sd(list_norm[["c"]])
mean_and_sd(list_norm[["d"]])
```

To write 4 rows at once can click shift option then highlight the rows that i want to type on

But imagine there are more than 4 samples, what if had 10000s of samples then would be very difficult. Can create a for loop for this 

## Use a for loop 

Create an output list, and run a for loop 

```{r}
#we are creating an output with 4 spots in it 

output = vector("list", length = 4)

for (i in 1:4) {
  
  output[[i]] = mean_and_sd(list_norm[[i]])
  
}
```

Got a list of normals, there are 4 lists and for any one of those i want to apply mean and sd. and now i am creating an output list to catch each of the outputs separately 

if list_norm had 1000 entries to it can still work (rather than keeping track of each index myself and typing it all out)
