---
title: "20241024_Writing_Functions"
output: github_document
date: "2024-10-24"
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
```

rvest to do some data importing 

Functions to use when using the same basic code more than two times (iteration)
Functions like mapping in maths (have an input then have a described operation and have an output)

Scoping: if in function there is some variable that you haven't defined then the function will scope your entire R environment to see whether you had previously defined this variable (but generally this is something that you do not want)
want to make sure that everything that you need for the function to work is defined inside the function (since work environemnt can change)
Function should be self contained and is everything that this function needs defined in the function 

Writing a function that is clear is really important 

## Writing my first function 

Standardising is frequent in statistic (subtract mean divide standard deviation)

As an example, here is a z-score computation 

```{r}
x_vec = rnorm(n = 25, mean = 10, sd = 3.5)

(x_vec - mean(x_vec)) / sd(x_vec)
```

Next write a function that calculates the z-score 

```{r}
z_scores = function(x) {
  
  z = (x - mean(x)) / sd(x)
  
  return(z)
}

z_scores(x = x_vec)
```

parenthesizes () is where define argument, {} to define the operation
  Can see () defining what my inout is and {} what my output should be 

Inside the function not re-using the x-vector 
Inside function i have the argument x and then outside the function i have defined that the x argument is the x-vector
  Can change the argument into something else and the function will operate the same 
  
Can create a new x vector 

Kinds of things that want to double check 
Does this always work

```{r, error = TRUE}
z_scores(x = 3)
```
Cannot work since there is no sd if there is only one value 

```{r, error = TRUE}
z_scores("my name is jeff")
```
Cannot work since trying to take the mean of words not numbers 


So these are checks and we can put this into the function 
Conditional execution checks
```{r}
z_scores = function(x) {
  
  if(!is.numeric(x)) {
    stop("x needs to be numeric")
  }
  
  if(length(x) < 5) {
    stop("you need at least 5 numbers to compute the z score")
  }
  
  z = (x - mean(x)) / sd(x)
  
  return(z)
}

z_scores(x = x_vec)
```

```{r, error = TRUE}
z_scores(x = 3)
```

```{r, error = TRUE}
z_scores("my name is jeff")
```

Now these can work

Helpful that conditional execution exists 
If the input is a character vector i want this to happen but if it is numeric i want this to happen (compute a plot or a table based on data input)

This is relatively simple (one input one output)



