---
title: "20241031_Simulation"
output: github_document
date: "2024-11-04"
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
```

Why simulations?
Central limit theorem, drawing independent samples and samples from a population eventually will average of the sample mean will look closer to population mean
repeated sampling (statistical analysis) does not happen in reality (data expensive to collect) but it can happen on a computer (can see what really happens under repeated sampling)

Gives us a way to check whether statistical approach does the thing that it claims to do 

Simulations are natural in the context of iterations 

Write a function that
- defines data generating mechanisms
- draw a sample
- analyse the sample 
- repeated object of interest (mean, regression etc)

use loop to repeat 

can inspect and understand properties of statistical analysis 
what happens when change sample size, what happens when change something in study design 

## Let's simulate something

sample size has default true mean 3 and true sd 4
```{r}
sim_mean_sd = function(samp_size, mu = 3, sigma = 4) {
  
  sim_data = 
    tibble(
      x = rnorm(n = samp_size, mean = mu, sd = sigma) #generates data from a normal dis with this sample size and this mean
    )
  
  sim_data %>% #then taking that data sample, we know the truth of the population
    summarise(
      mean = mean(x),
      sd = sd(x)
    ) #returns sample mean and sd
}

```

Repeating this run (go out into the population, draw sample size of 30, then give me sample mean and sample sd)

I have a function and i can simulate by running this line
```{r}
sim_mean_sd(30)
```
mean is around 3 and sd is around 4 like the true population 
This is the process that we are going to be more rigorous about 

Suppose running this over and over again, what would the distribution of the mean should look like?

(drawing from normal dis population, sample mean should be centered around true mean and sd of sample mean is the population sd divided by sample size); establish by statistics


## Let's stimualte a lot 

Not running same line of code over and over again, but doing it formally 

Start with a for loop
Need to have the output then have the for loop itself

```{r}
output = vector("list", length = 100)

for (i in 1:100) {
  
  output[[i]] = sim_mean_sd(samp_size = 30)
  
} #for each iteration from 1 to 100 i need to do something (run sim_mean_sdw with sample size 30)

#Then need to save the save the output of i

#more manageable if bind rows together 

bind_rows(output)
```

Looking at this:
I have gone out 100 times, each time taking a sample of 30 from a population with a true mean of 3 and sd of 4 
Given each sample compute the mean and sd for each sample 

## Let's use a loop function 

In the for loop previously there is an output loop, but there was no input list. 

In the case where we had an input this (for this input do this for this input do this), we used map 
Each input in the input list applied to function to get an output list 

Here different, we just want to do the same thing over and over again --> rerun 
rerun(some number of times, some thing)

```{r}
sim_results = 
  rerun(100, sim_mean_sd(samp_size = 30)) %>% 
  bind_rows()
```

## Let's take a look at the 100 sample mean and sd results that we got 

wrote a function, use rerun, and binding rows 

Can start to do data manipulation to this 

```{r}
sim_results %>% 
  ggplot(aes(x = mean)) + geom_density()

sim_results %>% 
  summarise(
    avg_samp_mean = mean(mean),
    sd_samp_mean = sd(mean)
  )
```
This shows the distribution of the sample mean across the 100 experiments 
(centered around 3 which makes sense as the true population mean we set was 3)

And the sample sd should be able the population sd / sample size 30 ~ 0.7

```{r}
sim_results %>% 
  ggplot(aes(x = sd)) + geom_density()
```

dis of sample sd is about 4 but unlike the dis of sample mean the dis is not normally dis, there is a heavier tail 
Through simulation we can learn more about this dis, about what the shape of the dis of sample sd looks like in repeated sampling 









